#This R markdown script plots the figures as they appear in Bijl et al., in review in ESSD, using the data frames as stored in <https://github.com/bijlpeter83/DINOSTRAT.git>. This markdown is to be used freely, but please cite the corresponding paper in ESSD when using the data.

```{r}
library(readxl)
library(caTools)
library(ggspatial)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
library(ggplot2)
library(tidyverse)
library(geoscale)
library(jcolors)
library(rgeos)
library(patchwork)
library(RColorBrewer)

theme_set(theme_bw())
```

#loading sites into R, and selecting variables of interest to plot

```{r}
locations <- read_csv("https://raw.githubusercontent.com/bijlpeter83/DINOSTRAT/main/data/Paleolatitude.csv")
sites <- select(locations, location, long, lat, tier)
sites <- distinct(sites)
sites$tier <- as.character(sites$tier)
```
#Selecting variables of interest for plotting, and to make sure the variables have the right format:
```{r}
paleolat <- select(locations, Source, Geography, location, Age, Paleolatitude, lower, upper, interpolated, Stable, tier)
paleolat$tier <- as.character(paleolat$tier)
paleolat$Paleolatitude <- as.numeric(paleolat$Paleolatitude)
paleolat$lower <- as.numeric(paleolat$lower)
paleolat$upper <- as.numeric(paleolat$upper)
paleolat
```
#adding the locations of surface sediment samples:
```{r}
modernst <- read_csv("https://raw.githubusercontent.com/bijlpeter83/DINOSTRAT/main/data/modernst.csv")
```
#selecting input for plotting maps:
```{r}
world <- ne_countries(scale = "medium", returnclass = "sf")
class(world)
```
#plot the global map of sites used in this study and the detailed map of sites in Europe:

```{r}
pW <- ggplot()+
  geom_sf(data = world, fill="darkgrey")+
  coord_sf()+
  geom_point(modernst, mapping=aes(x=long, y=lat), color="black", size = 0.1)+
  geom_point(sites, mapping = aes(x=long, y=lat, color = tier), size=1)+
scale_x_continuous(limits=c(-180, 180), breaks = seq(-180, 180, 30), expand = expansion(mult=0, add=0))+
  scale_y_continuous(expand = expansion(mult=0, add=0), limits=c(-90, 90), breaks = seq(-90, 90, 30))+
  scale_color_jcolors()+
  theme(axis.title = element_blank(), axis.text = element_blank(), panel.background = element_rect(fill = "lightgrey"))+
  xlab("Longitude") + ylab("Latitude") +
    ggtitle(paste0("               Global map of calibrated dinocyst records ", "(", length(unique(sites$location)), " sites)"))

pE <- ggplot(data = world) +
  geom_sf(fill="darkgrey")+
  geom_point(modernst, mapping=aes(x=long, y=lat), color="black", size = 0.1)+
  geom_point(sites, mapping = aes(x=long, y=lat, color = tier), size=1)+
  scale_color_jcolors()+
  coord_sf(xlim = c(-30, 50), ylim = c(35, 70))+
  theme(panel.background = element_rect(fill="lightgrey"))+
  xlab("Longitude") + ylab("Latitude")+
  ggtitle("Detailed map of European calibrated dinoflagellate cyst records")
```
#plot the latitudinal and longitudinal histograms:
```{r}
plo <- ggplot()+
  geom_histogram(sites, mapping = aes(x=long), bins = 72, fill = "red", alpha=0.5)+
scale_x_continuous(limits=c(-180, 180), breaks = seq(-180, 180, 30), expand = expansion(mult=0, add=0))+
  scale_y_continuous(expand = expansion(mult=0, add=0), position = "right")+
  theme_bw()+
  theme(panel.grid=element_blank(), rect = element_blank(), axis.title.x = element_blank())+
  coord_fixed()

pla <- ggplot()+
  geom_histogram(sites, mapping=aes(y=lat), bins = 36, fill = "red", alpha=0.5, orientation = "y")+
  scale_y_continuous(limits=c(-90, 90), breaks = seq(-90, 90, 30), expand = expansion(mult=0, add=0))+
  scale_x_continuous(expand = expansion(mult=0, add=0), position = "top")+
  theme_bw()+
  theme(panel.grid=element_blank(), rect = element_blank(), axis.title.y = element_blank())+
  coord_fixed()
#ggsave("latitude distribution.pdf")

pwo <- pW +inset_element(plo, left=-0.005, right=1.074, bottom = -0.13, top = 0.3, align_to = 'plot')+
  inset_element(pla, left=-0.225, right=0.3, bottom=-0.01, top=1.137, align_to = 'plot')
```
#I pasted these plots together in Illustrator, to make Figure 2

#To plot Figure 3:
```{r}
clrs <- read_csv("https://raw.githubusercontent.com/bijlpeter83/DINOSTRAT/main/data/CGMW_ICS_colour.csv")
clrs <- filter(clrs, Cyan>=0)
clrs <- select(clrs, 'Long List (formatted)', 'RGB', 'Red', 'Green', 'Blue')
clrs <- rename(clrs, 'stage'= 'Long List (formatted)')

modernst <- read_csv("https://raw.githubusercontent.com/bijlpeter83/DINOSTRAT/main/data/stages.csv")
stages <- right_join(stg, clrs, by= "stage")
typcol <- select(stages, stage, Red, Green, Blue)
typcol <- transmute(typcol, stage, RGB= rgb(Red, Green, Blue, maxColorValue = 255))
stgcol <- setNames(as.character(typcol$RGB), as.character(typcol$stage))
```
#To plot the paleolatitudes of site through time:
```{r}
ggplot(paleolat)+
  geom_ribbon(mapping=aes(x=Age, ymin=lower, ymax=upper, group=location), alpha=0.1)+
  geom_line(mapping=aes(x=Age, y=Paleolatitude, color=tier, group=location, linewidth = Stable), orientation = "x", lineend = "round")+
  scale_linewidth_discrete(range=c(0.4, 0.8))+
  scale_x_continuous(name="Age (Ma; GTS2020)", breaks = seq(0, 250, 25), minor_breaks = seq(0, 250, 5), labels = seq(0, 250, 25), expand = expansion(mult=0, add=0))+
  scale_y_continuous(name = "Paleolatitude", breaks = seq(-90, 90, 30), minor_breaks = seq(-90, 90, 10), labels = seq(-90, 90, 30), expand = expansion(mult=0, add=0))+
  scale_color_jcolors()+
  geom_rect(stages, mapping=aes(xmin=bage, xmax=tage, ymin=-100, ymax=-80), color = "Black", fill=stgcol)+
  geom_text(stages, mapping=aes(label=stage, y=-90, x=(tage+(bage-tage)/2)), size = 1.6, angle=90)

#ggsave("paleolat.pdf")
```
#to load the dinocyst fossil and modern events into R:
```{r}
sp_dat <- read_csv("https://raw.githubusercontent.com/bijlpeter83/DINOSTRAT/main/data/Dinoevents_FEB_23.csv")
sp_dat <- unite(sp_dat, Family, Subfamily, col="Family", sep=", ", na.rm=TRUE, remove = TRUE)
sp_dat <- unite(sp_dat, genus, species, subspecies, col = "taxa", sep= " ", na.rm=TRUE, remove=FALSE)

modernsp <- read_csv("https://raw.githubusercontent.com/bijlpeter83/DINOSTRAT/main/data/modernsp.csv")
modernsp$subspecies <- as.character(modernsp$subspecies)
modernsp <- unite(modernsp, Family, Subfamily, col="Family", sep=", ", na.rm=TRUE, remove = TRUE)
modernsp <- unite(modernsp, genus, species, subspecies, col = "taxa", sep= " ", na.rm=TRUE, remove=FALSE)
modernsp$subspecies <- as.character(modernsp$subspecies)
modernsp$'original entry' <- as.character(modernsp$'original entry')

sp_dat1 <- bind_rows(sp_dat, modernsp)
FO_sp <- filter(sp_dat1, FO_LO=="FO")
LO_sp <- filter(sp_dat1, FO_LO=="LO")
```
#These will filter the data to oldest FOs and youngest LOs of taxa:
```{r}
Hist_FO <- filter(FO_sp%>%group_by(taxa), Age==max(Age))
Hist_LO <- filter(LO_sp%>%group_by(taxa), Age==min(Age))
Hist <- bind_rows(Hist_FO, Hist_LO)
Hist_GF <- filter(FO_sp%>%group_by(genus), Age==max(Age))
Hist_GL <- filter(LO_sp%>%group_by(genus), Age==min(Age))
Hist_G <- bind_rows(Hist_GF, Hist_GL)
```
#Histograms per family:
```{r}
for (i in unique(sp_dat1$Family)) {
 pdf(file=paste("distribution Family ", i, ".pdf", sep=""))
  plot(ggplot()+
  geom_histogram(filter(Hist_FO, Family==i), mapping=aes(y=Age, x=..count..), bins = 125, fill="blue", alpha=0.5)+
  geom_histogram(filter(Hist_LO, Family==i), mapping=aes(y=Age, x=-..count..), bins = 125, fill="red3", alpha=0.5)+
  geom_histogram(filter(Hist_GF, Family==i), mapping=aes(y=Age, x=..count..), bins = 125, fill="blue")+
  geom_histogram(filter(Hist_GL, Family==i), mapping=aes(y=Age, x=-..count..), bins = 125, fill="red3")+
    geom_segment(filter(Hist, Family==i), mapping=aes(x=0, xend=0, y=max(Age), yend=min(Age)), linewidth=0.25)+
  scale_y_continuous(limits=c(250, 0), breaks = seq(0, 250, 25), expand = expansion(mult=0, add=0), trans="reverse")+
  scale_x_continuous(limits=c(-25, 25), expand = expansion(mult=0, add=0))+
  theme(aspect.ratio = 3, panel.grid = element_blank(), rect = element_blank())+
    ggtitle(paste0("Family ", i))
  ) 
  dev.off()
}
```
#Histogram of FOs and LOs of all taxa:
```{r}
ggplot()+
  geom_histogram(Hist_FO, mapping=aes(x=Age, y=..count..), bins = 125, fill="blue", alpha=0.5)+
  geom_histogram(Hist_LO, mapping=aes(x=Age, y=-..count..), bins = 125, fill="red3", alpha=0.5)+
  geom_histogram(Hist_GF, mapping=aes(x=Age, y=..count..), bins = 125, fill="blue")+
  geom_histogram(Hist_GL, mapping=aes(x=Age, y=-..count..), bins = 125, fill="red3")+
  scale_x_continuous(limits=c(0, 250), breaks = seq(0, 250, 25))+
  scale_y_continuous(limits=c(-80, 80))+
  theme(aspect.ratio = 0.5)
#ggsave("hist_age2.pdf")
```
#select the latitudinally synchronous FOs:
```{r}
Fsp <- c(
  "Achomosphaera andalousiensis", "Adnatosphaeridium vittanum", "Aldorfia aldorfensis", "Alisocysta margarita", "Ambonosphaera calloviana", "Amiculosphaera umbracula", "Apectodinium augustum", "Apectodinium hyperacanthum", "Aprobolocysta eilema", "Areoligera gippingensis", "Areosphaeridium diktyoplokum", "Artemisiocysta cladodichotoma", "Ataxiodinium confusum", "Atopodinium polygonale", "Atopodinium prostatum",
  "Biorbifera johnwingii", "Barbatacysta pilosa", "Barssidinium evangelineae", "Batioladinium reticulatum",
  "Callaiosphaeridium trycherium", "Carpathodinium predae", "Carpatella cornuta", "Cerodinium pannuceum", "Cerodinium speciosum", "Cerodinium striatum", "Cerodinium medcalfii", "Charlesdowniea columna", "Charlesdowniea edwardsii", "Chiropteridium lobospinosum", "Chytroeisphaeridia chytroeides", "Chytroeisphaeridia hyalina", "Cleistosphaeridium polypetellum", "Cleistosphaeridium varispinum", "Corrudinium harlandii", "Cribroperidinium tenuiceras", "Ctenidodinium cornigera",
 "Danea californica", "Danea impages", "Dapcodinium priscum", "Dichadogonyaulax culmula", "Dinogymnium albertii", "Diphyes ficusoides", "Distatodinium biffii", "Dracodinium condylos", "Dracodinium pachydermum", "Dracodinium rhomboideum", "Dracodinium varielongitudum", 
  "Ectosphaeropsis burdigalensis", "Endoscrinium galeritum", "Escharisphaeridia pocockii", "Enneadocysta arcuata",
  "Glaphyrocysta divaricata", "Glaphyrocysta semitecta", "Gonyaulacysta jurassica", "Gonyaulacysta pectinigera", "Glossodinium dimorphum", "Glaphyrocysta pastielsii", "Gonyaulacysta centriconnata", "Gonyaulacysta eisenackii",
  "Hystrichodinium furcatum", "Hemiplacophora semilunifera", "Herendeenia postprojecta", "Heterosphaeridium difficile", "Homotryblium floripes", "Homotryblium tasmaniense", "Hystrichokolpoma spinosum",
  "Kisselevia insolens", 
  "Leptodinium ambiguum", "Liasidium variabile", "Limbodinium absidatum", "Litosphaeridium arundum", "Litosphaeridium siphoniphorum", 
  "Mancodinium semitabulatum", "Manumiella druggii", "Membranilarnacia tenella", "Mendicodinium brunneum", "Mendicodinium microscabratum", 
 "Nelchinopsis kostromiensis", 
  "Occisocysta balios", "Ovoidinium cinctum", "Oligokolpoma galeotti", 
  "Pareodinia ceratophora", "Pareodinia prolongata", "Pseudoceratium pelliferum", "Palaeohystrichophora infusorioides", "Parvocysta bullala", "Phallocysta eumekes", "Prolixosphaeridium parvispinum", 
  "Rigaudella aemula", "Rhombodinium perforatum", 
  "Sepispinula ancorifera", "Spiniferites manumii", "Susadinium scofoides", "Stephanelytron redcliffense", "Schematophora obscura", "Schematophora speciosa", "Scriniocassis priscus", "Scriniodinium crystallinum", "Scriniodinium dictyotum", "Scriniodinium pharo", "Selenopemphix armageddonensis", "Senoniasphaera inornata", "Sirmiodiniopsis orbis", "Sirmiodinium grossi", "Spiniferites manumii", "Spiniferites ramosus", "Stiphrosphaeridium dictyophorum", 
  "Trichodinium scarburghensis", "Tubotuberella dangeardii", 
  "Unipontidinium aquaeductum", 
  "Valvaeodinium spinosum", "Valensiella ovulum",
  "Wanaea digitata", "Wanaea fimbriata", "Wanaea indotata", "Wanaea thysanota", "Rhombodinium draco", "Rhombodinium perforatum", "Wetzeliella gochtii", "Wetzeliella meckelfeldensis", "Wetzeliella samlandica", "Wilsonidinium ornatum")
```
#and the latitudinally synchronous LOs:
```{r}
Lsp <- c(
  "Achomosphaera neptuni","Aldorfia aldorfensis", "Alisocysta margarita","Ambonosphaera calloviana", "Amiculosphaera umbracula", "Apteodinium granulatum", "Areoligera medusettiformis", "Areoligera tauloma", "Areosphaeridium diktyoplokum", "Atopodinium polygonale", "Atopodinium prostatum", "Apectodinium augustum", 
  "Batiacasphaera compta", "Biorbifera johnwingii", "Barssidinium graminosum", "Barssidinium pliocenicum", "Batiacasphaera minuta", "Batioladinium reticulatum", 
  "Carpatella cornuta", "Cerodinium diebelli", "Cerodinium striatum", "Cerodinium medcalfii", "Charlesdowniea columna", "Charlesdowniea edwardsii", "Cooksonidium capricornum", "Cordosphaeridium funiculatum", "Corrudinium devernalliae", "Corrudinium harlandii", "Cribroperidinium wetzelii", "Ctenidodinium continuum", "Cymosphaeridium validum",
 "Dapcodinium priscum", "Dichadogonyaulax culmula", "Dracodinium pachydermum", "Dracodinium rhomboideum", "Distatodinium biffi", "Dingodinium spinosum", "Dinogymnium albertii", 
  "Endoscrinium galeritum", "Eisenackia reticulata", "Enneadocysta multicornuta",
  "Fibradinium annetorpense", 
 "Glossodinium dimorphum", "Gonyaulacysta centriconnata", 
 "Hemiplacophora semilunifera", "Heteraulacacysta porosa",
 "Isabelidinium viborgense",
  "Kleithriasphaeridium corrugatum", "Kisselevia insolens", 
  "Labyrinthodinium truncatum", "Liasidium variabile", "Licracysta semicirculata", "Limbodinium absidatum", "Litosphaeridium arundum", "Litosphaeridium siphoniphorum", 
  "Mancodinium semitabulatum", "Manumiella druggii", "Membranilarnacia picena", "Membranilarnacia tenella", 
  "Odontochitina costata", "Operculodinium crassum",
  "Pareodinia prolongata", "Phoberocysta neocomica", "Pseudoceratium pelliferum", "Palaeoperidinium pyrophorum", "Phthanoperidinium comatum", "Palynodinium grallator", "Pseudoceratium pelliferum", 
  "Rhaetogonyaulax rhaetica", "Rhombodinium perforatum", 
 "Saturnodinium pansum", "Scriniocassis weberi", "Scriniocassis priscus", "Spiniferites manumii", "Spiniferites ramosus", "Stephanelytron redcliffense", "Susadinium scofoides", "Suessia swabiana", "Schematophora speciosa", "Scriniocassis priscus", "Senoniasphaera inornata", "Spiniferites manumii", "Stephodinium coronatum", 
  "Trichodinium scarburghensis", "Trithyrodinium evittii", "Trithyrodinium suspectum", 
  "Unipontidinium aquaeductum",
  "Valvaeodinium koessenium", 
  "Wanaea digitata", "Wanaea fimbriata", "Wanaea indotata", "Wanaea thysanota", "Wetzeliella gochtii", "Wetzeliella samlandica", "Wilsonidinium ornatum",  
  "Xenascus ceratoides")

FO_sp_s <- filter(FO_sp, taxa %in% Fsp)
LO_sp_s <- filter(LO_sp, taxa %in% Lsp)

FO_sp_s2 <- filter(FO_sp_s, Family=="Gonyaulacaceae, Leptodinioideae")
LO_sp_s2 <- filter(LO_sp_s, Family=="Gonyaulacaceae, Leptodinioideae")

FO_sp_s2
LO_sp_s2

FOn <- unique(FO_sp_s2$taxa)
LOn <- unique(LO_sp_s2$taxa)
```
#This code will plot a PDF Age over paleolatitude of dinocyst events per family of dinocysts, consistent with Figure 5-28. Plots are stored on your computer (search for "Family_"), and will not show here.
```{r}
for (i in unique(sp_dat1$Family)) {
  
  pdf(file=paste("Family_", i, ".pdf", sep=""))
  
  plot(ggplot()+
  geom_line(paleolat, mapping=aes(x=Age, y=Paleolatitude, group=location), size=0.4, orientation = "x", color = "grey", alpha=0.5)+
  geom_point(filter(FO_sp, Family==i), mapping=aes(x=Age, y=pal_lat), size=0.6, shape = 17)+
  geom_line(filter(FO_sp, Family==i), mapping=aes(x=Age, y=pal_lat, group=taxa), size=0.2, orientation = "y")+
  geom_point(filter(LO_sp, Family==i), mapping=aes(x=Age, y=pal_lat), size=0.6, shape = 19)+
  geom_line(filter(LO_sp, Family==i), mapping=aes(x=Age, y=pal_lat, group=taxa), size=0.2, linetype="longdash" , orientation = "y")+
   geom_line(filter(FO_sp_s, Family==i), mapping=aes(x=Age, y=pal_lat, group=taxa, color=taxa), size=0.6, orientation = "y")+
  geom_line(filter(LO_sp_s, Family==i), mapping=aes(x=Age, y=pal_lat, group=taxa, color=taxa), size=0.6, linetype="longdash", orientation = "y")+ 
    geom_point(filter(FO_sp_s, Family==i), mapping=aes(x=Age, y=pal_lat, group=taxa, color=taxa), size=0.6, shape = 17)+
    geom_point(filter(LO_sp_s, Family==i), mapping=aes(x=Age, y=pal_lat, group=taxa, color=taxa), size=0.6, shape = 19)+ 
  scale_x_continuous(name="Age (Ma; GTS2020)", breaks = seq(0, 250, 25), minor_breaks = seq(0, 250, 5), labels = seq(0, 250, 25), expand = expansion(mult=0, add=0))+
  scale_y_continuous(name = "Paleolatitude", breaks = seq(-90, 90, 30), minor_breaks = seq(-90, 90, 10), labels = seq(-90, 90, 30), expand = expansion(mult=0, add=0))+
  ggtitle(paste0("Family ", i))+
    coord_fixed(ratio=1)+
    theme_bw()+
    theme(legend.position = "bottom", legend.text = element_text(size=7, face="italic"))+
    geom_rect(stages, mapping=aes(xmin=bage, xmax=tage, ymin=-100, ymax=-80), color = "Black", fill=stgcol)+
  geom_text(stages, mapping=aes(label=stage, y=-90, x=(tage+(bage-tage)/2)), size = 1.6, angle=90)
  )
  
  dev.off()
  
}
```
#this code will plot PDFs of modern species as in Figure 4 and in Supplement 2. Plots will be saved on your computer, and will not show here:
```{r}
for(i in unique(modernsp$taxa)){
 
  pdf(file=paste("modern_", i, ".pdf", sep=""))
  
  plot(ggplot()+
  geom_line(paleolat, mapping=aes(x=Age, y=Paleolatitude, group=location), orientation = "x", color = "grey", alpha=0.5)+
  geom_point(filter(modernsp, genus ==i), mapping=aes(x=Age, y=pal_lat), size=0.2, color = "red")+
  geom_point(filter(FO_sp, taxa==i), mapping=aes(x=Age, y=pal_lat), size=0.2, color = "blue")+
  geom_line(filter(FO_sp, taxa==i), mapping=aes(x=Age, y=pal_lat, group=taxa), size=0.2, color = "blue", orientation = "y")+
  geom_point(filter(LO_sp, taxa==i), mapping=aes(x=Age, y=pal_lat), size=0.2, color = "red")+
  geom_line(filter(LO_sp, taxa==i), mapping=aes(x=Age, y=pal_lat, group=taxa), size=0.2, color = "red", orientation = "y")+
  scale_x_continuous(name="Age (Ma; GTS2020)", breaks = seq(0, 250, 10), minor_breaks = seq(0, 250, 5), expand = expansion(mult=0, add=1))+
  scale_y_continuous(name = "Paleolatitude", breaks = seq(-90, 90, 30), minor_breaks = seq(-90, 90, 10), labels = seq(-90, 90, 30), expand = expansion(mult=0, add=0))+
    theme_bw()+
  ggtitle(paste0("Species ", i))+
  geom_rect(stages, mapping=aes(xmin=bage, xmax=tage, ymin=-100, ymax=-80), color = "Black", fill=stgcol)+
  geom_text(stages, mapping=aes(label=stage, y=-90, x=(tage+(bage-tage)/2)), size = 1.6, angle=90)
  )

dev.off()  
  
}
```
#add modern lat and long to the paleodata:
```{r}
amodernst <- select(modernst, location, lat, long)
asites <- select(sites, location, lat, long)
alatlong <- bind_rows(amodernst, asites)
alatlong

sp_dat3 <- right_join(sp_dat1,alatlong, by='location')
sp_datm <- filter(sp_dat3, Age == 0)
sp_datp <- filter(sp_dat3, Age>0)

sp_datm
sp_datp
```
#this code will plot latitude over age, maps, and histograms over age PDFs per genus. Plots will be saved on your computer, and will not show here:
```{r}
for(i in unique(sp_dat3$genus)){
 
  pdf(file=paste("Distribution ", i, ".pdf", sep=""))
  
  plot(ggplot()+
  geom_line(paleolat, mapping=aes(x=Age, y=Paleolatitude, group=location), orientation = "x", color = "grey", alpha=0.5)+
  geom_point(filter(LO_sp, genus == i), mapping=aes(x=Age, y=pal_lat), size=0.2, color = "red")+
  geom_line(filter(LO_sp, genus == i), mapping=aes(x=Age, y=pal_lat, group=taxa), linewidth=0.2, color = "red", orientation = "y")+
  geom_point(filter(FO_sp, genus == i), mapping=aes(x=Age, y=pal_lat), size=0.2, color = "blue")+
  geom_line(filter(FO_sp, genus == i), mapping=aes(x=Age, y=pal_lat, group=taxa), linewidth=0.2, color = "blue", orientation = "y")+
  scale_x_continuous(name="Age (Ma; GTS2020)", breaks = seq(0, 250, 10), minor_breaks = seq(0, 250, 5), expand = expansion(mult=0, add=1))+
  scale_y_continuous(name = "Paleolatitude", breaks = seq(-90, 90, 30), minor_breaks = seq(-90, 90, 10), labels = seq(-90, 90, 30), expand = expansion(mult=0, add=0))+
  ggtitle(paste0(i))+
    geom_rect(stages, mapping=aes(xmin=bage, xmax=tage, ymin=-100, ymax=-80), color = "Black", fill=stgcol)+
  geom_text(stages, mapping=aes(label=stage, y=-90, x=(tage+(bage-tage)/2)), size = 1.6, angle=90)
  )
 
dev.off() 

}


  for(i in unique(sp_dat3$genus)){
 
  pdf(file=paste("Histogram ", i, ".pdf", sep=""))
    
  plot(ggplot()+
  geom_histogram(filter(Hist_FO, genus==i), mapping=aes(x=Age, y=..count..), bins = 125, fill="blue")+
  geom_histogram(filter(Hist_LO, genus==i), mapping=aes(x=Age, y=-..count..), bins = 125, fill="red3")+
    geom_hline(mapping=aes(yintercept=0))+
  scale_y_continuous(limits=c(-5, 3), breaks = seq(-5, 3, 1), expand = expansion(mult=0, add=0))+
    scale_x_continuous(limits=c(0,250), breaks = seq(0,250,10), minor_breaks = seq(0,250,5), expand = expansion(mult=0, add=1))+
  theme(aspect.ratio = 0.3, panel.grid = element_blank())+
    ggtitle(paste0(i))+
    geom_rect(stages, mapping=aes(xmin=bage, xmax=tage, ymin=-5, ymax=-3), color = "Black", fill=stgcol)+
  geom_text(stages, mapping=aes(label=stage, y=-4, x=(tage+(bage-tage)/2)), size = 1.6, angle=90)
  )
 
dev.off() 

  }
  

  for(i in unique(sp_dat3$genus)){
 
  pdf(file=paste("Map ", i, ".pdf", sep=""))
    
  plot(ggplot()+
    geom_sf(data = world, fill="darkgrey")+
  coord_sf()+
  geom_point(filter(sp_datm, genus == i), mapping=aes(x=long, y=lat), color="darkgreen", size=0.5)+
  geom_point(filter(sp_datp, genus == i), mapping = aes(x=long, y=lat), color = "green")+
scale_x_continuous(limits=c(-180, 180), breaks = seq(-180, 180, 30), expand = expansion(mult=0, add=0))+
  scale_y_continuous(expand = expansion(mult=0, add=0), limits=c(-90, 90), breaks = seq(-90, 90, 30))+
  theme(panel.background = element_rect(fill = "lightblue"))+
  xlab("Longitude")+
  ylab("Latitude")+
     ggtitle(paste0("Genus " , i)))
 
dev.off() 

  }
  
  
```
#this code will plot latitude over age, maps, and histograms over age PDFs per unique taxon. Plots will be saved on your computer, and will not show here:

```{r}
for(i in unique(sp_dat3$taxa)){
 
  pdf(file=paste("Distribution ", i, ".pdf", sep=""))
  
  plot(ggplot()+
  geom_line(paleolat, mapping=aes(x=Age, y=Paleolatitude, group=location), orientation = "x", color = "grey", alpha=0.5)+
  geom_point(filter(LO_sp, taxa == i), mapping=aes(x=Age, y=pal_lat), size=0.2, color = "red")+
  geom_line(filter(LO_sp, taxa == i), mapping=aes(x=Age, y=pal_lat, group=taxa), size=0.2, color = "red", orientation = "y")+
  geom_point(filter(FO_sp, taxa == i), mapping=aes(x=Age, y=pal_lat), size=0.2, color = "blue")+
  geom_line(filter(FO_sp, taxa == i), mapping=aes(x=Age, y=pal_lat, group=taxa), size=0.2, color = "blue", orientation = "y")+
  scale_x_continuous(name="Age (Ma; GTS2020)", breaks = seq(0, 250, 10), expand = expansion(mult=0, add=1))+
  scale_y_continuous(name = "Paleolatitude", breaks = seq(-90, 90, 30), minor_breaks = seq(-90, 90, 10), labels = seq(-90, 90, 30), expand = expansion(mult=0, add=0))+
  ggtitle(paste0(i))+
    geom_rect(stages, mapping=aes(xmin=bage, xmax=tage, ymin=-100, ymax=-80), color = "Black", fill=stgcol)+
  geom_text(stages, mapping=aes(label=stage, y=-90, x=(tage+(bage-tage)/2)), size = 1.6, angle=90)
  )
 
dev.off() 

}


  for(i in unique(sp_dat3$taxa)){
 
  pdf(file=paste("Histogram ", i, ".pdf", sep=""))
    
  plot(ggplot()+
  geom_histogram(filter(Hist_FO, taxa==i), mapping=aes(x=Age, y=..count..), bins = 125, fill="blue")+
  geom_histogram(filter(Hist_LO, taxa==i), mapping=aes(x=Age, y=-..count..), bins = 125, fill="red3")+
    geom_segment(filter(Hist, taxa==i), mapping=aes(x=max(Age), y=0, xend=min(Age), yend=0), linewidth=0.25)+
  scale_y_continuous(limits=c(-5, 3), breaks = seq(-5, 3, 1), expand = expansion(mult=0, add=0))+
    scale_x_continuous(limits=c(0,250), breaks = seq(0,250,10), expand = expansion(mult=0, add=1))+
  theme(aspect.ratio = 0.3, panel.grid = element_blank())+
    ggtitle(paste0(i))+
    geom_rect(stages, mapping=aes(xmin=bage, xmax=tage, ymin=-5, ymax=-3), color = "Black", fill=stgcol)+
  geom_text(stages, mapping=aes(label=stage, y=-4, x=(tage+(bage-tage)/2)), size = 1.6, angle=90)
  )
 
dev.off() 

  }
  

  for(i in unique(sp_dat3$taxa)){
 
  pdf(file=paste("Map ", i, ".pdf", sep=""))
    
  plot(ggplot()+
    geom_sf(data = world, fill="darkgrey")+
  coord_sf()+
  geom_point(filter(sp_datm, taxa == i), mapping=aes(x=long, y=lat), color="darkgreen", size = 0.5)+
  geom_point(filter(sp_datp, taxa == i), mapping = aes(x=long, y=lat), color = "green")+
scale_x_continuous(limits=c(-180, 180), breaks = seq(-180, 180, 30), expand = expansion(mult=0, add=0))+
  scale_y_continuous(expand = expansion(mult=0, add=0), limits=c(-90, 90), breaks = seq(-90, 90, 30))+
  theme(panel.background = element_rect(fill = "lightblue"))+
  xlab("Longitude")+
  ylab("Latitude")+
     ggtitle(paste0("Species " , i)))
 
dev.off() 

  }
  

```
#Plots stratigraphic ranges of species per site. The plots will be stored on your computer, and not show here.
```{r}
for(i in unique(sp_dat$location)){
# first determine how wide the plot should be
  basewidth <- 2 # in inches
  specwidth <- 0.1 # width per species band? approximately
  nspecies <- sp_dat %>%
    filter(location == i) %>%
    nrow()

  # then create the plotting space
  pdf(file=paste("Site_", i, ".pdf", sep=""), width = basewidth + specwidth * nspecies)

plot(ggplot(filter(sp_dat, location== i), mapping=aes(x=fct_reorder(taxa, Age, max, .desc=TRUE), y=Age))+
  geom_line()+
  geom_point(aes(shape=FO_LO), fill= "white")+
    scale_shape_manual(values = c(24, 25))+
    scale_y_continuous(trans="reverse")+
  theme_bw()+
  theme(axis.text.x = element_text(face= "italic", angle=90, hjust = 1, vjust=0.2, size = 4))+
    xlab("Species")+
    ylab("Age (Ma; GTS2020)")+
  ggtitle("Site ", i))

dev.off()

}
```